Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2013-10-09 paper/paper.tex
Constants parameter

\subsection{$l$: Program Length}
The first paramter we introduce is program length, denoted by $l$.
At each iteration we synthesise programs of length exactly $l$.
We start with $l = 1$ and increment $l$ by $1$ when we have determined
that no program of length $l$ can satisfy the specifcation.  When we do
successfully synthesise a program, we are guaranteed that it
is of minimal length since we have previously established that no
correct program of smaller length exists.

\subsection{$w$: Word Width}
It is often the case that a program which satisfies the specification will
continue to satisfy it regardless of the word width of the machine it is
implemented on.  For example, the program in Fig.~\ref{fig:bitvector-program}
isolates the least-significant bit of an 8-bit word, but it will also
isolate the least-significant bit of a 32-bit word or indeed of a word of
any size.

@@

\subsection{$c$: Number of Constants}
$\mathcal{L}$ is a SSA, three address instruction set.  Destination registers
are implicit and a fresh register exists for each instruction to write its
output to.  A natural way to encode $\mathcal{L}$ instructions is to have an
opcode and two operands.  The opcode selects which instruction type is being
executed and the operands specify which registers should be operated on.
We might allow the operands to be either a register (i.e. a program argument
or the result of a previous instruction), or an immediate constant.

Each opcode requires $\lceil \log_2 I \rceil$ bits to encode, where $I$ is the number
of instruction types in $\mathcal{L}$.  Each operand can be encoded using
$\log_2 w$ bits, where $w$ is the $\mathcal{L}$-machine word, plus one
bit to specify whether the operand is a register name or an immediate constant.
One instruction can therefore be encoded using $\lceil \log_2 I \rceil + 2w + 2$ bits.
For an $n$ instruction program, we need $$\lceil n \log_2 I \rceil + 2nw + 2n$$ bits to encode
the entire program.

If we instead limit the number of constants that can appear in the program,
our operands can be encoded using fewer bits.  For an $n$ instruction program
using $k$ constants and taking $a$ arguments as inputs, each operand can refer
to a program argument, the result of a previous instruction or a constant.
This can be encoded using $\lceil \log_2 (k+a+n-1) \rceil$ bits, which means each instruction
can be encoded in $\lceil \log_2 I \rceil + \lceil \log_2 (k + a + n - 1) \rceil$ and the full program
needs $$\lceil n \log_2 I \rceil + \lceil n \log_2 (k + a + n - 1) \rceil + kw$$ bits to encode.

As with program length we can progressively increase the number of constants in
our program.  We start by trying to synthesise a program with no constants, then
if that fails we try to synthesise using one constant and so on.

>>>>

\subsection{Program Length: $l$}
The first paramter we introduce is program length, denoted by $l$.
At each iteration we synthesise programs of length exactly $l$.
We start with $l = 1$ and increment $l$ by $1$ when we have determined
that no program of length $l$ can satisfy the specifcation.  When we do
successfully synthesise a program, we are guaranteed that it
is of minimal length since we have previously established that no
correct program of smaller length exists.

\subsection{Number of Constants: $c$}
Instructions in $\mathcal{L}$ take either one or two operands.
Since any instruction whose operands are all constants can always be
eliminated (since its result is a constant), we know that a program
of minimal length will not contain any instructions with two constant
operands.  Therefore the number of constants which can appear in
a minimal program of length $l$ is at most $2l$.  By minimising the number
of constants appearing in a program, we are able to use a particularly
efficient program encoding that speeds up the synthesis procedure
substantially.  The number of constants used in a program is the parameter $c$.

$\mathcal{L}$ is a SSA, three address instruction set.  Destination registers
are implicit and a fresh register exists for each instruction to write its
output to.  A na\"{\i}ve way to encode $\mathcal{L}$ instructions is to have an
opcode and two operands, where each operand is either a register (i.e. a program argument
or the result of a previous instruction), or an immediate constant.

In this encoding, each opcode requires $\lceil \log_2 I \rceil$ bits to encode, where $I$ is the number
of instruction types in $\mathcal{L}$.  Each operand can be encoded using
$\log_2 w$ bits, where $w$ is the $\mathcal{L}$-machine word, plus one
bit to specify whether the operand is a register name or an immediate constant.
One instruction can therefore be encoded using $\lceil \log_2 I \rceil + 2w + 2$ bits.
For an $n$ instruction program, we need $$\lceil n \log_2 I \rceil + 2nw + 2n$$ bits to encode
the entire program.

If we instead limit the number of constants that can appear in the program,
our operands can be encoded using fewer bits.  For an $n$ instruction program
using $c$ constants and taking $a$ arguments as inputs, each operand can refer
to a program argument, the result of a previous instruction or a constant.
This can be encoded using $\lceil \log_2 (c+a+n-1) \rceil$ bits, which means each instruction
can be encoded in $\lceil \log_2 I \rceil + \lceil \log_2 (c + a + n - 1) \rceil$ and the full program
needs $$\lceil n \log_2 I \rceil + \lceil n \log_2 (c + a + n - 1) \rceil + cw$$ bits to encode.

As with program length we can progressively increase the number of constants in
our program.  We start by trying to synthesise a program with no constants, then
if that fails we try to synthesise using one constant and so on until we reach
$c = l$.

\subsection{Word Width: $w$}
It is often the case that a program which satisfies the specification will
continue to satisfy it regardless of the word width of the machine it is
implemented on.  For example, the program in Fig.~\ref{fig:bitvector-program}
isolates the least-significant bit of an 8-bit word, but it will also
isolate the least-significant bit of a 32-bit word or indeed of a word of
any size.

@@
